---
title: "Study Fleet Manuscript Pull and Analysis"
author: "Andy Jones"
date: Date & Time - `r format(Sys.time(), '%Y %B %d')`
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r global-options, include=FALSE}
knitr::opts_chunk$set(fig.path='Figs/',echo=TRUE, warning=FALSE, message=FALSE)

```


##Loading packages
Loading the many packages that are used somewhere in this script. This script has evolved to do many things and many of these are only need in one or two places.

```{r Loading packages}
#loading needed packages
library(tidyverse)
library(lubridate)
library(MASS)
library(sf)
library(gt)
library(pals)
library(scales)
library(mapdata)
library(patchwork)
library(ggnewscale)
library(stars)
library(raster)

```

##Load pulled data from file
```{r Loading a local version of the data}
#Data saved as an RData file
#Last pulled on 12/15/2021
load("C:/Users/andrew.jones/Desktop/Study_Fleet_Manuscript.RData")

```

Getting into the analysis with some summaries

# Table 1

Number of vessels, trips, and efforts by year and reporting type (contracted study Fleet ‘STFLT’ or other voluntary haul-level reporting Northeast Cooperative Research Program ‘NCRP’). The general trend in each category is of rapid increase, however the voluntary records do not begin until 2011. Vessels reporting at the haul-level that are not part of the contracted Study Fleet but were participating in another NEFSC program that mandated haul-level reporting (such as pilot electronic monitoring programs) are grouped under the NCRP category.

```{r Making data summaries}
#Summarizing the numbers for the study fleet and other haul-level reporting
trip_counts <- crpp_trip %>% mutate(YEAR=year(SAIL_DATE_GMT)) %>% dplyr::select(YEAR,SOURCE,TRIP_ID) %>% distinct() %>% group_by(YEAR,SOURCE) %>% tally() %>% pivot_wider(names_from = YEAR,values_from = n)

ves_counts <- crpp_trip %>% mutate(YEAR=year(SAIL_DATE_GMT)) %>% dplyr::select(YEAR,SOURCE,VESSEL_PERMIT_NUM) %>% distinct() %>% group_by(YEAR,SOURCE) %>% tally() %>% pivot_wider(names_from = YEAR,values_from = n)

effort_counts <- crpp_trip %>% mutate(YEAR=year(SAIL_DATE_GMT)) %>% dplyr::select(TRIP_ID,YEAR) %>% left_join(., crpp_effort %>% mutate(effort_id=paste(TRIP_ID,EFFORT_NUM))) %>% dplyr::select(YEAR,SOURCE,effort_id) %>% distinct() %>% group_by(YEAR,SOURCE) %>% tally() %>% pivot_wider(names_from = YEAR,values_from = n) %>% drop_na(SOURCE)

gear_counts <- crpp_trip %>% mutate(YEAR=year(SAIL_DATE_GMT)) %>% dplyr::select(TRIP_ID,YEAR) %>% left_join(., crpp_effort %>% mutate(effort_id=paste(TRIP_ID,EFFORT_NUM))) %>% dplyr::select(YEAR,GC_GEAR_CODE,TRIP_ID) %>% distinct() %>% 
  mutate(GC_GEAR_CODE=fct_lump_n(GC_GEAR_CODE,n=1)) %>%
  group_by(YEAR,GC_GEAR_CODE) %>% tally() %>% pivot_wider(names_from = YEAR,values_from = n) %>% drop_na(GC_GEAR_CODE)

```

```{r A table summary of study fleet data}
#Making two tables from these summaries
bind_rows(tibble(Metric='Number of Vessels',ves_counts),tibble(Metric='Number of Trips',trip_counts),tibble(Metric='Number of Efforts',effort_counts)) %>% dplyr::select(-`2006`,-`2021`) %>% replace(is.na(.), 0) %>% group_by(Metric) %>% gt() %>% tab_header(title = "Metrics of Study Fleet Participation", subtitle ="From 2007 - 2020")


```

# Table 2
The number of trips by gear type and year. Generally, otter trawl gear has been the dominant category of gear used by vessels in the program. Other gears have become more common in recent years suggesting that the ELB is capable of supporting these fleets.

```{r A summary of gear types in the study fleet data}
#Gear type summaries
gear_counts %>% dplyr::select(-`2006`,-`2021`) %>% replace(is.na(.), 0) %>%
  mutate(GC_GEAR_CODE = as.character(GC_GEAR_CODE)) %>%
  mutate(GC_GEAR_CODE=case_when(GC_GEAR_CODE=="092OTF" ~ 'Otter trawl fish',TRUE ~ GC_GEAR_CODE)) %>%
  rename('Gear type'=GC_GEAR_CODE) %>%
  gt() %>% 
  tab_header(title = "Study Fleet Participation by Gear Type",
             subtitle ="From 2007 - 2020")

```

# Figure 1

###ADD IMAGE###

# Figure 2
The number of trips that were collected by the ELB system. Haul-by-haul reporting (high-resolution) trips are shown in purple, and trip reporting (lower resolution) in yellow. This shows the expansion of the ELB for regulatory and scientific purposes.

```{r A figure of trip counts, fig.height=4, fig.width=6}
#Plotting out trip count figure
#For Study Fleet data
trip_list %>%  mutate(YEAR=year(SAIL_DATE_LCL)) %>% 
  group_by(YEAR,REPORT_SOURCE) %>% tally() %>%
  filter(YEAR > 2006, YEAR < 2021) %>%
  mutate(REPORT_SOURCE2=case_when(REPORT_SOURCE=='HBH'~'Haul-level',TRUE ~ 'Trip-level')) %>%
  #pivot_wider(names_from = YEAR,values_from = n)
  ggplot(aes(x=YEAR,y=n,color=REPORT_SOURCE2)) + geom_point(size=4,shape=1) + geom_path(size=2) +
  labs(x='Year',y='Number of trips',color='Report type') +
  scale_x_continuous(breaks=c(2007:2020)) + theme_bw() +
  scale_color_manual(values=c('#352A87','#FBCD2D')) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```

#Figure 3
The breakdown of Study Fleet number of vessels, trips, and hauls by target species through time. The trends through time for each level of the data are shown in panels A, B, and C. The annual proportions of vessels, trips, and hauls are shown in panels D, E, and F.

```{r Looking at the study fleet summaries by target species, fig.height=9, fig.width=9}
#Breaking out records by target (trips/)
target_haul_trend <- plot_data_comb_ves %>%
  mutate(target=as.character(target)) %>%
  mutate(target=case_when(target=='LOLIGO SQUID' ~ 'LONGFIN SQUID',
                          target=='ILLEX SQUID'~'SHORTFIN SQUID',TRUE ~ target)) %>%
  filter(source=='crb') %>%
  drop_na(target) %>%
  mutate(target=fct_lump_n(target,n=10)) %>%
  dplyr::select(haul_id,YEAR,target) %>%
  distinct() %>%
  group_by(YEAR,target) %>% tally() %>%
  filter(YEAR > 2006, YEAR < 2021) %>%
  ggplot(aes(x=YEAR,y=n,colour=target)) + geom_point(size=2,shape=1) + geom_path(size=1) + 
  labs(x='Year',y='Number of hauls',color='Target species') +
  scale_x_continuous(breaks=c(2007:2020)) + 
  theme_bw() +
  theme(legend.position = 'None') +
  scale_colour_manual(values=as.vector(parula(11))) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

target_haul_prop <- plot_data_comb_ves %>%
  mutate(target=as.character(target)) %>%
  mutate(target=case_when(target=='LOLIGO SQUID' ~ 'LONGFIN SQUID',
                          target=='ILLEX SQUID'~'SHORTFIN SQUID',TRUE ~ target)) %>%
  filter(source=='crb') %>%
  drop_na(target) %>%
  mutate(target=fct_lump_n(target,n=10)) %>%
  dplyr::select(haul_id,YEAR,target) %>%
  distinct() %>%
  group_by(YEAR,target) %>% tally() %>%
  filter(YEAR > 2006, YEAR < 2021) %>%
  ggplot(aes(x=YEAR,y=n,fill=target)) + geom_col(position='fill') +
  labs(x='Year',y='Percentage of hauls',fill='Target species') +
  scale_x_continuous(breaks=c(2007:2020)) + theme_bw() +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_manual(values=as.vector(parula(11))) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

target_trip_trend <- plot_data_comb_ves %>%
  mutate(target=as.character(target)) %>%
  mutate(target=case_when(target=='LOLIGO SQUID' ~ 'LONGFIN SQUID',
                          target=='ILLEX SQUID'~'SHORTFIN SQUID',TRUE ~ target)) %>%
  filter(source=='crb') %>%
  drop_na(target) %>%
  mutate(target=fct_lump_n(target,n=10)) %>%
  dplyr::select(trip_id,YEAR,target) %>%
  distinct() %>%
  group_by(YEAR,target) %>% tally() %>%
  filter(YEAR > 2006, YEAR < 2021) %>%
  ggplot(aes(x=YEAR,y=n,colour=target)) + geom_point(size=2,shape=1) + geom_path(size=1) + 
  labs(x='Year',y='Number of trips',color='Target species') +
  scale_x_continuous(breaks=c(2007:2020)) +
    theme_bw() +
  theme(legend.position = 'None') +
  scale_colour_manual(values=as.vector(parula(11))) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

target_trip_prop <- plot_data_comb_ves %>%
  mutate(target=as.character(target)) %>%
  mutate(target=case_when(target=='LOLIGO SQUID' ~ 'LONGFIN SQUID',
                          target=='ILLEX SQUID'~'SHORTFIN SQUID',TRUE ~ target)) %>%
  filter(source=='crb') %>%
  drop_na(target) %>%
  mutate(target=fct_lump_n(target,n=10)) %>%
  dplyr::select(trip_id,YEAR,target) %>%
  distinct() %>%
  group_by(YEAR,target) %>% tally() %>%
  filter(YEAR > 2006, YEAR < 2021) %>%
  ggplot(aes(x=YEAR,y=n,fill=target)) + geom_col(position='fill') +
  labs(x='Year',y='Percentage of trips',fill='Target species') +
  scale_x_continuous(breaks=c(2007:2020)) + theme_bw() +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_manual(values=as.vector(parula(11))) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

target_vessel_trend <- plot_data_comb_ves %>%
  mutate(target=as.character(target)) %>%
  mutate(target=case_when(target=='LOLIGO SQUID' ~ 'LONGFIN SQUID',target=='ILLEX SQUID'~'SHORTFIN SQUID',TRUE ~ target)) %>%
  filter(source=='crb') %>%
  drop_na(target) %>%
  mutate(target=fct_lump_n(target,n=10)) %>%
  dplyr::select(permit,YEAR,target) %>%
  distinct() %>%
  group_by(YEAR,target) %>% tally() %>%
  filter(YEAR > 2006, YEAR < 2021) %>%
  ggplot(aes(x=YEAR,y=n,colour=target)) + geom_point(size=2,shape=1) + geom_path(size=1) + 
  labs(x='Year',y='Percentage of vessels',color='Target species') +
  scale_x_continuous(breaks=c(2007:2020)) +
  theme_bw() +
  theme(legend.position = 'None') +
  scale_colour_manual(values=as.vector(parula(11))) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

target_vessel_prop <- plot_data_comb_ves %>%
  mutate(target=as.character(target)) %>%
  mutate(target=case_when(target=='LOLIGO SQUID' ~ 'LONGFIN SQUID',
                          target=='ILLEX SQUID'~'SHORTFIN SQUID',TRUE ~ target)) %>%
  filter(source=='crb') %>%
  drop_na(target) %>%
  mutate(target=fct_lump_n(target,n=10)) %>%
  dplyr::select(permit,YEAR,target) %>%
  distinct() %>%
  group_by(YEAR,target) %>% tally() %>%
  filter(YEAR > 2006, YEAR < 2021) %>%
  ggplot(aes(x=YEAR,y=n,fill=target)) + geom_col(position='fill') +
  labs(x='Year',y='Proiportion of vessels',fill='Target species') +
  scale_x_continuous(breaks=c(2007:2020)) + theme_bw() +
  scale_fill_manual(values=as.vector(parula(11))) +
  scale_y_continuous(labels = scales::percent_format()) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))


((target_vessel_trend + target_trip_trend + target_haul_trend) / (target_vessel_prop + target_trip_prop + target_haul_prop)) +
  plot_layout(guides = "collect") + plot_annotation(tag_levels = 'A') + theme(legend.position = 'bottom')


```

# Figure 4
Trends in annual trips (A) and hauls (B) for the Study Fleet program are shown in blue. For comparison the annual number of trips and hauls from trawl vessels is shown from another large monitoring program in the region, the fishery observer programs. The goals and sampling schemes of these programs are quite different, but together they represent a large amount of fishery dependent data from the region. Of note is the dip in total trips in the observer program in 2020, a result of a pause in operations for the COVID-19 pandemic.

```{r Plots comparing the number of records in the study fleet and observer programs, fig.height=7, fig.width=5}
#Plotting out a comparison to the observer programs (trips and hauls)
a <- plot_data_comb_ves %>%
  dplyr::select(trip_id,YEAR,source) %>%
  distinct() %>%
   group_by(YEAR,source) %>% tally() %>%
  filter(YEAR > 2006, YEAR < 2021) %>%
  mutate(source2=case_when(source=='obs'~'Observer',TRUE ~ 'Study Fleet')) %>%
  #pivot_wider(names_from = YEAR,values_from = n)
  ggplot(aes(x=YEAR,y=n,color=source2)) + geom_point(size=4,shape=1) + geom_path(size=2) +
  labs(x='Year',y='Number of trips',color='Data source') +
  scale_x_continuous(breaks=c(2007:2020)) + theme_bw() +
   scale_color_manual(values=c('#352A87','#FBCD2D')) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

b <- plot_data_comb_ves %>%
  dplyr::select(haul_id,YEAR,source) %>%
  distinct() %>%
   group_by(YEAR,source) %>% tally() %>%
  filter(YEAR > 2006, YEAR < 2021) %>%
  mutate(source2=case_when(source=='obs'~'Observer',TRUE ~ 'Study Fleet')) %>%
  #pivot_wider(names_from = YEAR,values_from = n)
  ggplot(aes(x=YEAR,y=n,color=source2)) + geom_point(size=4,shape=1) + geom_path(size=2) +
  labs(x='Year',y='Number of hauls',color='Data source') +
  scale_x_continuous(breaks=c(2007:2020)) + theme_bw() +
  scale_color_manual(values=c('#352A87','#FBCD2D')) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

library(patchwork)
a/b + plot_annotation(tag_levels = 'A')

```

# Figure 5
Maps showing the spatial distribution of Study Fleet trawl records for 2010 - 2020 gridded into 5-minute squares. In panel A, the total number of hauls are shown. For comparison the distribution of observer records for trawl gear are shown in panel B. The data collected by the programs spans similar geographic ranges, however the Study Fleet data set includes more records from southern New England and fewer records from the Gulf of Maine. Panel C shows the total number of trips by Study Fleet vessels, indicating a high degree of overlap between trip and haul level record totals. Panel D shows the total catch by Study Fleet vessels (lbs), again showing a higher degree of similarity to the distributions of trips and hauls. In all panels cells with fewer than three unique permits (records from fewer than three vessels) have been omitted.

```{r Making rasters of data and mapping out records, fig.height=9, fig.width=9}
#Making more maps that remove the cells with fewer than three permits
#Subsetting data down to just a single gear type (otter trawl)
plot_data_comb_map <- plot_data_comb_ves %>% 
  filter(str_sub(gear_code_vtr,1,2)=='OT') %>% drop_na('start_lon','start_lat')

#Pulling in bathy info and setting map boundaries
#Depths are in meters
atl <- marmap::getNOAA.bathy(-76.5,-65, 34.5, 45, res = 1, keep=TRUE)
lons = c(-75.5, -66)
lats = c(35.5, 44)

#Making an object to show the coastlines
reg = map_data("world2Hires")
reg = subset(reg, region %in% c('Canada', 'USA'))
reg$long = (360 - reg$long)*-1

#Making a raster grid
boxsize <- 10
xmn <- -77
xmx <- -57
ymn <- 30
ymx <- 50
bmin <- 3
nrow <- (ymx-ymn)*60/boxsize
ncol <- (xmx-xmn)*60/boxsize
crs <- CRS("+proj=longlat +datum=NAD83 +no_defs +ellps=GRS80 +towgs84=0,0,0")

#Setting the raser box size
r = raster(nrow=(ymx-ymn)*60/boxsize
	           , ncol= (xmx-xmn)*60/boxsize
	           , xmn = xmn
	           , xmx = xmx
	           , ymn = ymn
	           , ymx = ymx
	           , crs = crs) 

#Setting up the coordinates and CRS in the data
coordinates(plot_data_comb_map) <- c('start_lon','start_lat')
plot_data_comb_map@proj4string = CRS("+proj=longlat +datum=NAD83 +no_defs +ellps=GRS80 +towgs84=0,0,0")

####
#Making rasters of the study fleet data
crb_plot_data_comb_map_raster_weight <- rasterize(plot_data_comb_map %>% st_as_sf() %>% filter(source=='crb'), r, 'TOT_CATCH',fun=sum)
crb_plot_data_comb_map_raster_permits <- rasterize(plot_data_comb_map %>% st_as_sf() %>% filter(source=='crb'), r, 'permit', fun=function(x, ...) length(unique(x)))
crb_plot_data_comb_map_raster_efforts <- rasterize(plot_data_comb_map %>% st_as_sf() %>% filter(source=='crb'), r, 'haul_id', fun=function(x, ...) length(unique(x)))
crb_plot_data_comb_map_raster_trips <- rasterize(plot_data_comb_map %>% st_as_sf() %>% filter(source=='crb'), r, 'trip_id', fun=function(x, ...) length(unique(x)))

#Stacking the study fleet rasters
crb_stack <- stack(crb_plot_data_comb_map_raster_efforts,crb_plot_data_comb_map_raster_permits,crb_plot_data_comb_map_raster_weight,crb_plot_data_comb_map_raster_trips) %>% st_as_stars(values=layer) %>% st_as_sf() %>% filter(layer.2>3) %>% drop_na(layer.1)

#####
#Making rasters for the observer data 
obs_plot_data_comb_map_raster_weight <- rasterize(plot_data_comb_map %>% st_as_sf() %>% filter(source=='obs'), r, 'TOT_CATCH',fun=sum)
obs_plot_data_comb_map_raster_permits <- rasterize(plot_data_comb_map %>% st_as_sf() %>% filter(source=='obs'), r, 'permit', fun=function(x, ...) length(unique(x)))
obs_plot_data_comb_map_raster_efforts <- rasterize(plot_data_comb_map %>% st_as_sf() %>% filter(source=='obs'), r, 'haul_id', fun=function(x, ...) length(unique(x)))

#Stacking the observer rasters
obs_stack <- stack(obs_plot_data_comb_map_raster_efforts,obs_plot_data_comb_map_raster_permits) %>% st_as_stars(values=layer) %>% st_as_sf() %>% filter(layer.2>3) %>% drop_na(layer.1)

#Making an observer plot
observer <- ggplot() +
  scale_fill_gradientn(colours = brewer.purples(10),guide='colourbar',trans='log10') +
  geom_sf(data=obs_stack,aes(fill=layer.1)) +
  geom_polygon(data = reg, aes(x=long, y = lat, group = group),fill='grey60',colour='dark gray') +
  geom_contour(data = atl,
                aes(x=x, y=y, z=z),
                breaks=c(-50,-100,-1000),
                size=c(0.5),
                colour="black",alpha=0.5) +
  coord_sf(xlim = lons, ylim = lats,crs="+proj=longlat +datum=NAD83") +
    labs(x='Longitude',y='Latitude',fill='Total\nhauls') + 
  theme_bw() +
  theme(legend.position = c(0.85, 0.25))

#Making a study fleet plots
study_fleet <- ggplot() +
  scale_fill_gradientn(colours = brewer.blues(10),guide='colourbar',trans='log10') +
  geom_sf(data=crb_stack,aes(fill=layer.1)) +
  geom_polygon(data = reg, aes(x=long, y = lat, group = group),fill='grey60',colour='dark gray') +
  geom_contour(data = atl,
                aes(x=x, y=y, z=z),
                breaks=c(-50,-100,-1000),
                size=c(0.5),
                colour="black",alpha=0.5) +
  coord_sf(xlim = lons, ylim = lats,crs="+proj=longlat +datum=WGS84") +
    labs(x='Longitude',y='Latitude',fill='Total\nhauls') + 
  theme_bw() +
  theme(legend.position = c(0.85, 0.25))

study_fleet_w <- ggplot() +
  scale_fill_gradientn(colours = brewer.blues(10),guide='colourbar',trans='log10') +
  geom_sf(data=crb_stack,aes(fill=layer.3)) +
  geom_polygon(data = reg, aes(x=long, y = lat, group = group),fill='grey60',colour='dark gray') +
  geom_contour(data = atl,
                aes(x=x, y=y, z=z),
                breaks=c(-50,-100,-1000),
                size=c(0.5),
                colour="black",alpha=0.5) +
  coord_sf(xlim = lons, ylim = lats,crs="+proj=longlat +datum=WGS84") +
    labs(x='Longitude',y='Latitude',fill='Total\nweight (lbs)') + 
  theme_bw() +
  theme(legend.position = c(0.85, 0.25))

study_fleet_t <- ggplot() +
  scale_fill_gradientn(colours = brewer.blues(10),guide='colourbar',trans='log10') +
  geom_sf(data=crb_stack,aes(fill=layer.4)) +
  geom_polygon(data = reg, aes(x=long, y = lat, group = group),fill='grey60',colour='dark gray') +
  geom_contour(data = atl,
                aes(x=x, y=y, z=z),
                breaks=c(-50,-100,-1000),
                size=c(0.5),
                colour="black",alpha=0.5) +
  coord_sf(xlim = lons, ylim = lats,crs="+proj=longlat +datum=WGS84") +
    labs(x='Longitude',y='Latitude',fill='Total\ntrips') + 
  theme_bw() +
  theme(legend.position = c(0.85, 0.25))


#Putting the plots together
((study_fleet + observer) + plot_annotation(tag_levels = 'A')) / (study_fleet_t + study_fleet_w) + plot_annotation(tag_levels = 'A')


```

# Figure 6
The proportion of each species' regional landings reported through Study Fleet (purple), relative to the trend in total commercial landings of the species (yellow). Here six species of high commercial interest are shown. The trend for most species is of an increasing proportion in the Study Fleet data set relative to the whole fishery. However, as in the case of Atlantic cod this is partially driven by a decreasing trend in total catch.

```{r Comparisons of study fleet landings to total landings, fig.height=7, fig.width=10}
#Extracting the VTR data

#This runs into memory problems
#Making a sum of landings for the region for specific species
# dealer_sum <- do.call(rbind.data.frame,cfders_spp) %>% dplyr::select(SPECIES_ITIS,YEAR,SPPLNDLB) %>%
#   filter(SPECIES_ITIS %in% c('172735','082372','082521','169182','164744','172908','172877','164712')) %>%
#   mutate(YEAR=as.numeric(YEAR)) %>% group_by(YEAR,SPECIES_ITIS) %>% 
#   dplyr::summarise(SPPLNDLB=sum(SPPLNDLB,na.rm=TRUE))


dealer_sum_2007 <- cfders_spp[[1]] %>% dplyr::select(SPECIES_ITIS,YEAR,SPPLNDLB) %>%
  filter(SPECIES_ITIS %in% c('172735','082372','082521','169182','164744','172908','172877','164712')) %>%
  mutate(YEAR=as.numeric(YEAR)) %>% group_by(YEAR,SPECIES_ITIS) %>% 
  dplyr::summarise(SPPLNDLB=sum(SPPLNDLB,na.rm=TRUE))

dealer_sum_2007 <- cfders_spp[[2]] %>% dplyr::select(SPECIES_ITIS,YEAR,SPPLNDLB) %>%
  filter(SPECIES_ITIS %in% c('172735','082372','082521','169182','164744','172908','172877','164712')) %>%
  mutate(YEAR=as.numeric(YEAR)) %>% group_by(YEAR,SPECIES_ITIS) %>% 
  dplyr::summarise(SPPLNDLB=sum(SPPLNDLB,na.rm=TRUE))

dealer_sum_2008 <- cfders_spp[[3]] %>% dplyr::select(SPECIES_ITIS,YEAR,SPPLNDLB) %>%
  filter(SPECIES_ITIS %in% c('172735','082372','082521','169182','164744','172908','172877','164712')) %>%
  mutate(YEAR=as.numeric(YEAR)) %>% group_by(YEAR,SPECIES_ITIS) %>% 
  dplyr::summarise(SPPLNDLB=sum(SPPLNDLB,na.rm=TRUE))

dealer_sum_2009 <- cfders_spp[[4]] %>% dplyr::select(SPECIES_ITIS,YEAR,SPPLNDLB) %>%
  filter(SPECIES_ITIS %in% c('172735','082372','082521','169182','164744','172908','172877','164712')) %>%
  mutate(YEAR=as.numeric(YEAR)) %>% group_by(YEAR,SPECIES_ITIS) %>% 
  dplyr::summarise(SPPLNDLB=sum(SPPLNDLB,na.rm=TRUE))

dealer_sum_2010 <- cfders_spp[[1]] %>% dplyr::select(SPECIES_ITIS,YEAR,SPPLNDLB) %>%
  filter(SPECIES_ITIS %in% c('172735','082372','082521','169182','164744','172908','172877','164712')) %>%
  mutate(YEAR=as.numeric(YEAR)) %>% group_by(YEAR,SPECIES_ITIS) %>% 
  dplyr::summarise(SPPLNDLB=sum(SPPLNDLB,na.rm=TRUE))

dealer_sum_2011 <- cfders_spp[[5]] %>% dplyr::select(SPECIES_ITIS,YEAR,SPPLNDLB) %>%
  filter(SPECIES_ITIS %in% c('172735','082372','082521','169182','164744','172908','172877','164712')) %>%
  mutate(YEAR=as.numeric(YEAR)) %>% group_by(YEAR,SPECIES_ITIS) %>% 
  dplyr::summarise(SPPLNDLB=sum(SPPLNDLB,na.rm=TRUE))

dealer_sum_2012 <- cfders_spp[[6]] %>% dplyr::select(SPECIES_ITIS,YEAR,SPPLNDLB) %>%
  filter(SPECIES_ITIS %in% c('172735','082372','082521','169182','164744','172908','172877','164712')) %>%
  mutate(YEAR=as.numeric(YEAR)) %>% group_by(YEAR,SPECIES_ITIS) %>% 
  dplyr::summarise(SPPLNDLB=sum(SPPLNDLB,na.rm=TRUE))

dealer_sum_2013 <- cfders_spp[[7]] %>% dplyr::select(SPECIES_ITIS,YEAR,SPPLNDLB) %>%
  filter(SPECIES_ITIS %in% c('172735','082372','082521','169182','164744','172908','172877','164712')) %>%
  mutate(YEAR=as.numeric(YEAR)) %>% group_by(YEAR,SPECIES_ITIS) %>% 
  dplyr::summarise(SPPLNDLB=sum(SPPLNDLB,na.rm=TRUE))

dealer_sum_2014 <- cfders_spp[[8]] %>% dplyr::select(SPECIES_ITIS,YEAR,SPPLNDLB) %>%
  filter(SPECIES_ITIS %in% c('172735','082372','082521','169182','164744','172908','172877','164712')) %>%
  mutate(YEAR=as.numeric(YEAR)) %>% group_by(YEAR,SPECIES_ITIS) %>% 
  dplyr::summarise(SPPLNDLB=sum(SPPLNDLB,na.rm=TRUE))

dealer_sum_2015 <- cfders_spp[[9]] %>% dplyr::select(SPECIES_ITIS,YEAR,SPPLNDLB) %>%
  filter(SPECIES_ITIS %in% c('172735','082372','082521','169182','164744','172908','172877','164712')) %>%
  mutate(YEAR=as.numeric(YEAR)) %>% group_by(YEAR,SPECIES_ITIS) %>% 
  dplyr::summarise(SPPLNDLB=sum(SPPLNDLB,na.rm=TRUE))

dealer_sum_2016 <- cfders_spp[[10]] %>% dplyr::select(SPECIES_ITIS,YEAR,SPPLNDLB) %>%
  filter(SPECIES_ITIS %in% c('172735','082372','082521','169182','164744','172908','172877','164712')) %>%
  mutate(YEAR=as.numeric(YEAR)) %>% group_by(YEAR,SPECIES_ITIS) %>% 
  dplyr::summarise(SPPLNDLB=sum(SPPLNDLB,na.rm=TRUE))

dealer_sum_2017 <- cfders_spp[[11]] %>% dplyr::select(SPECIES_ITIS,YEAR,SPPLNDLB) %>%
  filter(SPECIES_ITIS %in% c('172735','082372','082521','169182','164744','172908','172877','164712')) %>%
  mutate(YEAR=as.numeric(YEAR)) %>% group_by(YEAR,SPECIES_ITIS) %>% 
  dplyr::summarise(SPPLNDLB=sum(SPPLNDLB,na.rm=TRUE))

dealer_sum_2018 <- cfders_spp[[12]] %>% dplyr::select(SPECIES_ITIS,YEAR,SPPLNDLB) %>%
  filter(SPECIES_ITIS %in% c('172735','082372','082521','169182','164744','172908','172877','164712')) %>%
  mutate(YEAR=as.numeric(YEAR)) %>% group_by(YEAR,SPECIES_ITIS) %>% 
  dplyr::summarise(SPPLNDLB=sum(SPPLNDLB,na.rm=TRUE))

dealer_sum_2019 <- cfders_spp[[13]] %>% dplyr::select(SPECIES_ITIS,YEAR,SPPLNDLB) %>%
  filter(SPECIES_ITIS %in% c('172735','082372','082521','169182','164744','172908','172877','164712')) %>%
  mutate(YEAR=as.numeric(YEAR)) %>% group_by(YEAR,SPECIES_ITIS) %>% 
  dplyr::summarise(SPPLNDLB=sum(SPPLNDLB,na.rm=TRUE))

dealer_sum_2020 <- cfders_spp[[14]] %>% dplyr::select(SPECIES_ITIS,YEAR,SPPLNDLB) %>%
  filter(SPECIES_ITIS %in% c('172735','082372','082521','169182','164744','172908','172877','164712')) %>%
  mutate(YEAR=as.numeric(YEAR)) %>% group_by(YEAR,SPECIES_ITIS) %>% 
  dplyr::summarise(SPPLNDLB=sum(SPPLNDLB,na.rm=TRUE))

dealer_sum <- rbind(dealer_sum_2020,dealer_sum_2019,dealer_sum_2018,dealer_sum_2017,
                    dealer_sum_2016,dealer_sum_2015,dealer_sum_2014,dealer_sum_2013,
                    dealer_sum_2012,dealer_sum_2011,dealer_sum_2010,dealer_sum_2009,
                    dealer_sum_2008,dealer_sum_2007)

#Making a sum for the study fleet
study_fleet_sum <- pulled_data_edit_cf %>%
  filter(SPECIES_ITIS %in% c('172735','082372','082521','169182','164744','172908','172877','164712')) %>%
  group_by(YEAR,SPECIES_ITIS) %>% sample_frac(0.5) %>%
  dplyr::summarise(SUM_HAIL_AMOUNT_LB=sum(HAIL_AMOUNT_LB,na.rm=TRUE))

#Putting these time
dealer_sum %>% inner_join(.,study_fleet_sum) %>% mutate(prop=SUM_HAIL_AMOUNT_LB/SPPLNDLB) %>%
  mutate(SPECIES_ITIS=case_when(SPECIES_ITIS=='082521'~'82521',SPECIES_ITIS=='082372'~'82372',TRUE~SPECIES_ITIS)) %>%
  left_join(.,OBSPEC %>% 
              filter(SPECIES_ITIS %in% c('172735','82372','82521',
                                         '169182','164744','172908','172877','164712')) %>%
            dplyr::select(SPECIES_ITIS,COMNAME) %>% distinct()
            ) %>%
  filter(COMNAME!='SQUID EGGS, ATL LONG-FIN',COMNAME!='FLOUNDER, AMERICAN PLAICE') %>%
  group_by(COMNAME) %>%
  mutate(SPPLNDLB=rescale(SPPLNDLB)) %>%
  ggplot() +
  geom_point(aes(x=YEAR,y=prop,color='Prop. in SF'),colour='#352A87', size=3,shape=1) +
  geom_line(aes(x=YEAR,y=prop,color='Prop. in SF'),colour='#352A87', size=1.2) +
  geom_point(aes(x=YEAR,y=SPPLNDLB/5,color='Total trend'),colour='#FBCD2D', size=3) +
  geom_line(aes(x=YEAR,y=SPPLNDLB/5,color='Total trend'),colour='#FBCD2D', size=1.2) +
  labs(x='Year',y='Proportion of total species landings (closed)') +
  scale_x_continuous(breaks=c(2007:2020)) + theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  facet_wrap(~COMNAME) +
  scale_y_continuous(sec.axis = sec_axis(~.*5, name = "Total landings trend (open)")) +
  theme(axis.line.y.right = element_line(color = "#FBCD2D"),
       axis.ticks.y.right = element_line(color = "#FBCD2D"),
       #axis.text.y.right = element_text(color = "#FBCD2D"),
       #axis.title.y.right = element_text(color="#FBCD2D"),
       axis.line.y.left = element_line(color = "#352A87"),
       axis.ticks.y.left = element_line(color = "#352A87"),
       axis.text.y.left = element_text(color = "#352A87"))

```

#Figure 7
The trend in oceanographic records through time and across space. Panel A shows the density of sampling across space for all years sampled (2010-2019).  Cells with fewer than three unique permits (records from fewer than three vessels) have been omitted. Panel B shows the number of fishing efforts with co-registered oceanographic data from 2010 to 2020. Data from 2019 and 2020 (shaded gray) are still being processed and sums represent incomplete totals.

```{r Makimh GTE plots, fig.height=6, fig.width=10}
#Making a similar plot for GTE data
plot_data_comb_ves_gte_filter <- plot_data_comb_ves_gte %>% filter(GTE_DATA=='YES') %>% 
  filter(YEAR>2009,YEAR<2021) %>% drop_na('start_lon','start_lat')

coordinates(plot_data_comb_ves_gte_filter) <- c('start_lon','start_lat')
plot_data_comb_ves_gte_filter @proj4string = CRS("+proj=longlat +datum=NAD83 +no_defs +ellps=GRS80 +towgs84=0,0,0")

crb_plot_data_comb_raster_efforts_gte <- rasterize(plot_data_comb_ves_gte_filter, r, 'haul_id', fun=function(x, ...) length(unique(x)))
crb_plot_data_comb_raster_permits_gte <- rasterize(plot_data_comb_ves_gte_filter, r, 'permit', fun=function(x, ...) length(unique(x)))

crb_stack_gte  <- stack(crb_plot_data_comb_raster_efforts_gte,crb_plot_data_comb_raster_permits_gte) %>% st_as_stars(values=layer) %>% st_as_sf() %>% filter(layer.2>3) %>% drop_na(layer.1)

#Map of GTE data
map_gte_2 <- ggplot() +
  scale_fill_gradientn(colours = brewer.blues(10),guide='colourbar',trans='log10') +
  geom_sf(data=crb_stack_gte,aes(fill=layer.1)) +
  geom_polygon(data = reg, aes(x=long, y = lat, group = group),
               fill='grey60',colour='white') +
  geom_contour(data = atl,
                aes(x=x, y=y, z=z),
                breaks=c(-50,-100,-1000),
                size=c(0.5),
                colour="black",alpha=0.5) +
  coord_sf(xlim = lons, ylim = lats,crs="+proj=longlat +datum=WGS84") +
    labs(x='Longitude',y='Latitude',fill='Total\nhauls') + 
  theme_bw() +
  theme(legend.position = c(0.85, 0.25))

#Trends in GTE records by target species
gte_trends <- plot_data_comb_ves_gte %>% filter(GTE_DATA=='YES',YEAR<2020) %>%
  mutate(target=as.character(target)) %>%
  mutate(target=case_when(target=='LOLIGO SQUID' ~ 'LONGFIN SQUID',
                          target=='ILLEX'~'SHORTFIN SQUID',TRUE ~ target)) %>%
  mutate(target=fct_lump_n(target,n=10)) %>%
  drop_na(target) %>%
  dplyr::select(haul_id,YEAR,target) %>% distinct() %>% 
  group_by(YEAR,target) %>% tally() %>%
  filter(YEAR>2009) %>% ggplot(aes(x=YEAR,y=n,colour=target)) + 
  geom_vline(xintercept = 2019.5,size=20,alpha=0.5) +
  geom_point(size=2) + geom_line(size=1.2) +
    labs(x='Year',y='Number of oceanographic records',colour='Target species') +
  scale_x_continuous(breaks=c(2010:2020),limits=c(2010,2020)) + theme_bw() +
  scale_colour_manual(values=as.vector(parula(11))) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

  
map_gte_2 + gte_trends + plot_annotation(tag_levels = 'A')

```

# Appendix C
Participation in the haul-by-haul program between 2007 and 2020. The number of trips in each year for each vessel is shown with shading (darker colors indicate more trips in a given year). Vessels are broken out by the state which the vessel is associated with. Other includes vessels from NY, NH, and ME.

```{r Making plots of ind vessel participation, fig.height=9, fig.width=9}
#Bringing together participant data
vessels_of_interest <- pulled_data_edit_cf$VESSEL_PERMIT_NUM %>% unique()

#Making a count of trips per year per vessel
participation_data <- pulled_data_edit_cf %>% 
  dplyr::select(YEAR,VESSEL_NAME,VESSEL_PERMIT_NUM,TRIP_ID) %>% 
  mutate(VESSEL_NAME = toupper(VESSEL_NAME)) %>%
  distinct() %>% group_by(YEAR,VESSEL_NAME,VESSEL_PERMIT_NUM) %>% tally() %>%
  filter(YEAR  > 2006) %>% 
  mutate(year_sum=sum(n)) %>%
  ungroup() %>%
  group_by(VESSEL_NAME,VESSEL_PERMIT_NUM) %>% 
  mutate(sum=sum(n)) %>%
  mutate(min_year=min(YEAR)) %>%
  #filter(sum > 500) %>%
  arrange(-sum)

VESSEL_NUMBER <- participation_data %>% ungroup() %>% 
  arrange(year_sum) %>% dplyr::select(VESSEL_NAME,VESSEL_PERMIT_NUM) %>%
  distinct() %>% mutate(VESSEL_NUM=row_number())

VESSEL_PORT <- VESSEL_NUMBER %>%
  left_join(.,vessel_info %>% mutate(VP_NUM=as.character(VP_NUM)) %>%
              filter(VP_NUM %in% vessels_of_interest,AP_YEAR>2006) %>%
              ungroup() %>%
              dplyr::select(VES_NAME,VP_NUM,PPST) %>% distinct() %>% 
              group_by(VP_NUM,VES_NAME) %>% slice(1), 
            by=c("VESSEL_PERMIT_NUM"="VP_NUM","VESSEL_NAME"="VES_NAME"))


#This is the final plot of participation by state
participation_data %>% left_join(VESSEL_PORT) %>% filter(YEAR < 2021) %>%
  ungroup() %>%
  mutate(Port=fct_lump(PPST,3)) %>%
  drop_na() %>%
  ggplot(aes(x=YEAR,y=reorder(VESSEL_NAME,-min_year))) + 
  geom_tile(aes(fill=n),colour='Gray') +
  labs(x='Year',y='Vessels',fill='Number of trips') +
  scale_x_continuous(breaks=c(2007:2020)) + theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.75, hjust=1),
        axis.text.y = element_text(size=6),
        legend.position = 'bottom') +
  facet_wrap(~Port,scales='free_y') +
  scale_fill_gradient(low=as.vector(brewer.blues(3)[1]),
                      high=as.vector(brewer.blues(3)[3]))

```
